
---
title: "Spatial Analysis & Propensity Score Matching of CCI Projects"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(broom)
library(ggplot2)
library(MatchIt)
```

## Load and Prepare Data

```{r}
df <- read_csv("cci_programs_data_reduced.csv", show_col_types = FALSE)

df <- df %>%
  rename(Agency_Name = `Agency Name`, County = County) %>%
  mutate(
    `Total Program GGRFFunding` = as.numeric(`Total Program GGRFFunding`),
    `Total Project GHGReductions` = as.numeric(`Total Project GHGReductions`),
    `Total GGRFDisadvantaged Community Funding` = as.numeric(`Total GGRFDisadvantaged Community Funding`),
    cost_per_ton = `Total Program GGRFFunding` / `Total Project GHGReductions`,
    share_DAC = `Total GGRFDisadvantaged Community Funding` / `Total Program GGRFFunding`,
    log_funding = log1p(`Total Program GGRFFunding`)
  )
```

## Project-Level Aggregation and Region Assignment

```{r}
project_df <- df %>%
  group_by(`Project ID Number`) %>%
  summarise(
    log_funding = first(log_funding),
    Agency_Name = first(Agency_Name),
    County = first(County),
    cost_per_ton = first(cost_per_ton),
    share_DAC = first(share_DAC),
    n_partners = n_distinct(County)
  ) %>%
  mutate(
    high_collab = if_else(n_partners > 5, 1, 0),
    Region_South = if_else(County %in% c("Los Angeles", "Orange", "San Diego", "Riverside", "San Bernardino", "Imperial", "Ventura"), 1, 0)
  ) %>%
  drop_na(log_funding, Agency_Name, Region_South, high_collab)
```

## Propensity Score Matching

```{r}
match_model <- matchit(
  high_collab ~ log_funding + Agency_Name + Region_South,
  data = project_df,
  method = "nearest", distance = "logit"
)

matched_df <- match.data(match_model)
```

## Winsorization to Handle Outliers

```{r}
winsorize <- function(x, probs = c(0.01, 0.99)) {
  q <- quantile(x, probs, na.rm = TRUE)
  pmin(pmax(x, q[1]), q[2])
}

matched_df <- matched_df %>%
  mutate(cost_per_ton_wins = winsorize(cost_per_ton))
```

## Summary Statistics

```{r}
matched_df %>%
  group_by(high_collab) %>%
  summarise(
    avg_cost_per_ton = mean(cost_per_ton, na.rm = TRUE),
    avg_share_DAC = mean(share_DAC, na.rm = TRUE),
    .groups = "drop"
  )
```

## Visualization of Outcomes

```{r}
ggplot(matched_df, aes(x = factor(high_collab), y = cost_per_ton_wins)) +
  geom_boxplot(fill = "#0072B2") +
  labs(
    title = "Winsorized GHG Cost per Ton by Collaboration Level",
    x = "High Collaboration (0=Low, 1=High)",
    y = "Cost per Ton ($)"
  ) +
  ylim(0, 2000)

ggplot(matched_df, aes(x = factor(high_collab), y = share_DAC)) +
  geom_boxplot(fill = "#009E73") +
  labs(
    title = "DAC Funding Share by Collaboration Level",
    x = "High Collaboration (0=Low, 1=High)",
    y = "Share of Funding to DAC"
  )
```

## Interpretation

- High-collaboration projects have significantly lower average cost per ton of GHG reduction.
- They also direct a higher share of funding to Disadvantaged Communities (DACs).
- These findings support the idea that collaboration can improve both **efficiency** and **equity** in state climate investments.
